---
title: "Prework"
subtitle: "R/Bioconductor for Bioinformatics"
format: 
  html:
    code-fold: false
    code-tools: false
editor: source
---

Before attending the workshop please have the following installed and configured on your machine.

-   Recent version of **R**

-   Recent version of **RStudio**

-   Most recent release of the [**Bioconductor**](https://www.bioconductor.org/install/) and other packages used in courses

Install the latest release of R, then get the latest version of Bioconductor by starting R and entering the commands. 

```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.16")
```

- Ensure you can knit R markdown documents

    -   Open RStudio and create a new Rmarkdown document
    -   Save the document and check you are able to knit it.

## Install Bioconductor Packages
```{r, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```

Install specific packages, e.g., “GenomicFeatures” and “AnnotationDbi”, with


```{r, eval=FALSE}
BiocManager::install(c("GenomicFeatures", "AnnotationDbi"))
```

The install() function (in the BiocManager package) has arguments that change its default behavior; type ?install for further help.


[Posit cloud](https://posit.cloud) is free of charge for personal users, yet you need to [sign up](https://posit.cloud/plans/free) for a new user account and have internet connection.

## Packages 

```{r}
#| eval: false
install.packages('BiocManager')
BiocManager::install('methods')
BiocManager::install('rmarkdown')
BiocManager::install('igraph')
BiocManager::install('org.Hs.eg.db')
BiocManager::install('RMariaDB')
BiocManager::install('Rbowtie')
BiocManager::install('parallel')
BiocManager::install('QuasR')
BiocManager::install('BSgenome')
BiocManager::install('BSgenome.Mmusculus.UCSC.mm10')
BiocManager::install('BSgenome.Hsapiens.UCSC.hg19')
BiocManager::install('BSgenome.Hsapiens.UCSC.hg38')
BiocManager::install('Biostrings')
BiocManager::install('rtracklayer')
BiocManager::install('ggplot2')
BiocManager::install('GenomicRanges')
BiocManager::install('ShortRead')
BiocManager::install('Rsubread')
BiocManager::install('Rsamtools')
BiocManager::install('Rbowtie2')
BiocManager::install('stringi')
BiocManager::install('Rhisat2')
BiocManager::install('R.utils')
BiocManager::install('GenomicAlignments')
BiocManager::install('basecallQC')
BiocManager::install('TxDb.Mmusculus.UCSC.mm10.knownGene')
BiocManager::install('TxDb.Hsapiens.UCSC.hg19.knownGene')
BiocManager::install('Rfastp')
BiocManager::install('TxDb.Hsapiens.UCSC.hg38.knownGene')
BiocManager::install('knitr')
BiocManager::install('testthat')
BiocManager::install('yaml')
```


## Data

The datasets we use can be found here <https://github.com/omicscenter/RforBioinformatics>.

## Code

The R scripts used in part 1 and part 2 can be found here <https://github.com/omicscenter/RforBioinformatics>.

## Resources

Lecture notes <https://github.com/omicscenter/RforBioinformatics>

Lab notes <https://github.com/omicscenter/RforBioinformatics>
